---
title: "String Functions"
author: "Hauke Sonnenberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{String Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The following functions are returning or analysing strings.

### Function `appendSuffix()`

This function adds a suffix to all elements of a vector of character. It is 
mainly a shortcut to `paste0(values, suffix)` but allows to define values to be 
omitted:

```{r}
# Define a vector of character
values <- c("id", letters[1:5])

# Add a suffix to all elements except "id"
kwb.utils::appendSuffix(values, "_value", valuesToOmit = "id")
```

### Function collapsed()

The function call `collapsed(x, sep)` is a shortcut to 
`paste(x, collapse = sep)`. It allows for shorter code and improved readability.

```{r}
kwb.utils::collapsed(1:10)
kwb.utils::collapsed(1:10, ", ")
kwb.utils::collapsed(letters, "|")
```

### Function commaCollapsed()

The function call `commaCollapsed(x)` is a shortcut to 
`paste(x, collapse = ",")`. It allows for shorter code and improved readability.

```{r}
kwb.utils::commaCollapsed(1:10)
kwb.utils::commaCollapsed(letters)
```

### Function defaultLevels()

When given a vector of character, this function returns the sorted unique 
values of the input vector. When given a vector of numeric with all differences 
being a multiple of the `step` argument (defaulting to `1`), 
the `step`-incremented sequence of values between the smallest and the highest 
value is returned.

```{r}
# Vector of character -> The sorted unique values are returned 
kwb.utils::defaultLevels(c("melon", "apple", "melon", "pear", "apple"))

# The differences between the values are not all a multiple of the step
# -> Return sorted unique values
kwb.utils::defaultLevels(c(5, 8, 1), step = 2)

# The differences between the values are all a multiple of the step
# -> Return step-incremented sequence between min and max
kwb.utils::defaultLevels(c(1.5, 4.5, 10.5), step = 1.5)
```


### Function extractSubstring()
### Function frenchToAscii()
### Function hsCountInStr()
### Function hsQuoteChr()
### Function hsSafeName()
### Function hsSubstSpecChars()
### Function hsTrim()
### Function moveToFront()
### Function multiSubstitute()
### Function pairwise()
### Function removeExtension()
### Function removeSpaces()
### Function shorten()
### Function stringContains()
### Function stringEndsWith()
### Function stringList()
### Function stringStartsWith()
### Function stringToExpression()
### Function subExpressionMatches()
