---
title: "Non-Categorised Functions"
author: "Hauke Sonnenberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Non-Categorised Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Function `arglist()`

The main purpose of the package "kwb.utils" is to provide functions that allow
for writing cleaner and safer code. One main characteristic of clean code is
that it is free or at least almost free of duplication. This is a very hard goal 
to achieve! But we should try our best!

Have you ever seen something like the following?

```{r}
x <- seq(-pi, pi, pi/10)
par(mfrow = c(1, 2), mar = c(2, 2, 1, 1))
plot(x, sin(x), xlab = "x", ylab = "sin(x)", pch = 16, type = "b", cex = 0.5)
plot(x, cos(x), xlab = "x", ylab = "cos(x)", pch = 16, type = "b", cex = 0.5)
```

Do you spot the repetition? What if you want to change the plot character `pch`
or the `type` of the plot or the character extension factor `cex` in both
plots at the same time?

```{r}
# Define common arguments
common <- list(x = x, pch = 16, xlab = "x", type = "b", cex = 0.5)

# Call the plot function with the common arguments, extended by
# call-specific arguments
par(mfrow = c(1, 2), mar = c(2, 2, 1, 1))
do.call(plot, kwb.utils::arglist(common, y = sin(x), ylab = "sin(x)"))
do.call(plot, kwb.utils::arglist(common, y = cos(x), ylab = "cos(x)"))
```

The function `arglist()` merges the contents of one or more lists by allowing
to overwrite existing list elements. In the above example, the list of `common`
arguments is just extended. Elements in this list could also be overwritten,
such as the argument `cex` in the following:

```{r}
do.call(
  what = plot, 
  args = kwb.utils::arglist(common, y = sin(x), ylab = "sin(x)", cex = 1)
)
```

This pattern of calling a function with an extended argument list is so common 
that I wrote a function `callWith()` that allows to write the above shorter:

```{r}
kwb.utils::callWith(plot, common, y = sin(x), ylab = "sin(x)", cex = 1)
```

But this will be explained later...



### Function DIN.A4()
### Function expandGrid()
### Function makeUnique()
### Function randomValuesWithSum()
