---
title: "Overview"
author: "Hauke Sonnenberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6
)
```

Today is a good day to start documenting what this package (kwb.utils) can be
used for. 

## Intention

The main purpose of the package "kwb.utils" is to provide functions that allow
for writing cleaner and safer code. One main characteristic of clean code is
that it is free or at least almost free of duplication. This is a very hard goal
to achieve! But we should try our best!

I would like to start with a complete list of the names of the
functions that are contained in this package. But how can I get such a list?

## How to get a list of functions in a package?

In fact, I do not know (any more) how to achieve this. However, I know that this
package contains a function `assignPackageObjects()` that copies all the objects
that are contained in a package (including all its functions) to the global
environment. By inspecting the definition, i.e. the body, of this function, we
can learn how this function works. Therefore, just type the function's name, 
without parentheses, and press Enter:

```{r}
kwb.utils::assignPackageObjects
```

As the code reveals, the function loops through the values that are returned by
`ls(getNamespace(package))` where `package` is the name of the package. This is
exacly what we need. We can use this code to get the names of all objects in the
"kwb.utils" package. Let's store these names in the variable `object_names`: 

```{r}
object_names <- ls(getNamespace("kwb.utils"))

length(object_names)
```

Wow! There are `r length(object_names)` objects (functions) defined in this
package. So where should I start? 

To make it somehow scientific, let's have a look at the distribution of the
function names, grouped by their first letter:

```{r}
letter_distribution <- table(substr(sort(tolower(object_names)), 1, 1))
barplot(letter_distribution, horiz = TRUE, las = 1)
```

Even though not the highest peak, there are a lot of function names starting
with `h`. Among these are functions that are prefixed by `hs`. We could find the
names of these functions with the function `stringStartsWith()`, that is
provided in this package (see [String Functions](#string), below). As, since R
version 3.3.0, there is a function `startsWith()` in base R, we should use that
one instead:

```{r}
object_names[startsWith(object_names, "hs")]
```

The functions prefixed by `hs` are the oldest. When I started writing functions,
I thought that it is important to indicate that these functions are user-defined
functions, as opposed to base R functions. Reading the book `Clean Code` by
Robert C. Martin convinced me to use clean names that can be pronounced easily.

## Functions by Category

I tried to group all package functions per topic or per type of object that they
either work on or return. These groups are listed below and described in further
detail in one dedicated article per group.

### Functions Returning Logical {#logical}

These functions take different types of arguments as input but have all in
common that they return a vector of logical. Some of these functions are just a
shortcut to a rather long expression aiming at making the code more readable and
avoiding duplication, such as `allAreEqual()`, `allAreIdentical()`,
`almostEqual()`, `isNaOrEmpty()` or `isNullOrEmpty()`. Others do more complex
things, such as `matchesCriteria()` that evaluates a set of conditions on a data
frame, returning `TRUE` or `FALSE` depending on whether a row matches all
conditions or not. Another more complex function is `hsValidValue()` that checks
if a vector of character seems to contain numerical values.

For details, see [Functions Returning Logical](functions_logical.html).

### String Functions {#string}

The functions prefixed by `string` are wrappers around some simple `grep`
expressions and exist since a former trainee asked me to provide these
functions. 

For details, see [String Functions](functions_string.html).

### Input and Output Functions

For details, see [Input and Output Functions](functions_in_out.html).

### Mathematical or Statistical Functions

For details, see [Mathematical or Statistical Functions](functions_stats.html).

### Functions Working on Data Frames

For details, see [Functions Working on Data Frames](functions_data_frame.html).

### Functions Working on Lists

For details, see [Functions Working on Lists](functions_list.html).

### Array/Matrix Functions

For details, see [Array/Matrix Functions](functions_array.html).

### Vector Functions

For details, see [Vector Functions](functions_vector.html).

### General Object Functions

For details, see [General Object Functions](functions_object.html).

### Conversion Functions

For details, see [Conversion Functions](functions_conversion.html).

### Shortcut Functions to if-statements

For details, see [Shortcut Functions to if-statements](functions_if.html).

### Operating System Functions

For details, see [Operating System Functions](functions_system.html).

### Check Functions

For details, see [Check Functions](functions_check.html).: Functions that check for prerequisites

### Debug Functions

For details, see [Debug Functions](functions_debug.html).

### Compare Objects

For details, see [Compare Objects](functions_debug.html).

### Calling Functions

For details, see [Calling Functions](functions_call.html).

### Cryptographic Functions

For details, see [Cryptographic Functions](functions_crypto.html).

### Grammar/Dictionary Functions

For details, see [Grammar/Dictionary Functions](functions_dictionary.html).

### Non-Categorised Functions

For details, see [Non-Categorised Functions](functions_ungrouped.html).
