---
title: "Overview"
author: "Hauke Sonnenberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6
)
```

Today is a good day to start documenting what this package (kwb.utils) can be
used for. I would like to start with a complete list of the names of the
functions that are contained in this package. But how can I get such a list?

## How to get an overview of functions in a package?

In fact, I do not know (any more) how to achieve this. However, I know that this
package contains a function `assignPackageObjects()` that copies all the objects
that are contained in a package (including all its functions) to the global
environment. By inspecting the definition, i.e. the body, of this function, we
can learn how this function works. Therefore, just type the function's name, 
without parentheses, and press Enter:

```{r}
kwb.utils::assignPackageObjects
```

As the code reveals, the function loops through the values that are returned by
`ls(getNamespace(package))` where `package` is the name of the package. This is
exacly what we need. We can use this code to get the names of all objects in the
"kwb.utils" package. Let's store these names in the variable `object_names`: 

```{r}
object_names <- ls(getNamespace("kwb.utils"))

length(object_names)
```

Wow! There are `r length(object_names)` objects (functions) defined in this
package. So where should I start? 

Let's have a look at the first six functions:

```{r}
head(object_names)
```

* [Functions Returning Logical](functions_logical.html)

* [String Functions](functions_string.html)

* [Input and Output Functions](functions_in_out.html)

* [Mathematical or Statistical Functions](functions_stats.html)

* [Functions Working on Data Frames](functions_data_frame.html)

* [Functions Working on Lists](functions_list.html)

* [Array/Matrix Functions](functions_array.html)

* [Vector Functions](functions_vector.html)

* [General Object Functions](functions_object.html)

* [Conversion Functions](functions_conversion.html)

* [Shortcut Functions to if-statements](functions_if.html)

* [Operating System Functions](functions_system.html)

* [Check Functions](functions_check.html): Functions that check for 
  prerequisites

* [Debug Functions](functions_debug.html)

* [Compare Objects](functions_debug.html)

* [Calling Functions](functions_call.html)

* [Cryptographic Functions](functions_crypto.html)

* [Grammar/Dictionary Functions](functions_dictionary.html)

* [Non-Categorised Functions](functions_ungrouped.html)
