---
title: "Overview"
author: "Hauke Sonnenberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6
)
```

Today is a good day to start documenting what this package (kwb.utils) is good
for. I would like to start with a complete list of the names of the functions
that are contained in this package. But how can I get such a list? 

## How to get an overview of functions in a package?

In fact, I do not know (any more) how to achieve this. However, I know that this
package contains a function `assignPackageObjects()` that copies all the objects
that are contained in a package (including all its functions) to the global
environment. By inspecting the definition, i.e. the body, of this function, we
can learn how this function works. Therefore, just type the function's name, 
without parentheses, and press Enter:

```{r}
kwb.utils::assignPackageObjects
```

As the code reveals, the function loops through the values that are returned by
`ls(getNamespace(package))` where `package` is the name of the package. This is
exacly what we need. We can use this code to get the names of all objects in the
"kwb.utils" package. Let's store these names in the variable `object_names`: 

```{r}
object_names <- ls(getNamespace("kwb.utils"))

length(object_names)
```

Wow! There are `r length(object_names)` objects (functions) defined in this
package. So where should I start? 

Let's have a look at the first six functions:

```{r}
head(object_names)
```

## Grouped Functions

`addClass()`: 
See article [General Object Functions](functions_object.html).

`addRowWithName()`: 
See article [Functions Working on Data Frames](functions_data_frame.html).

`addSuffixToColumns()`:
See article [Functions Working on Data Frames](functions_data_frame.html).

`allAreEqual()`:
See article [Functions Returning Logical](functions_logical.html).

`allAreIdentical()`:
See article [Functions Returning Logical](functions_logical.html).

`almostEqual()`:
See article [Functions Returning Logical](functions_logical.html).

`appendSuffix()`:
See article [String-Related Functions](functions_string.html).

`asColumnList()`:
See article [Conversion Functions](functions_conversion.html).

`asNoFactorDataFrame`:
See article [Functions Working on Data Frames](functions_data_frame.html).

`asRowList()`:
See article [Conversion Functions](functions_conversion.html).

`assertRowsAndColumns()`:
See article [Check Functions](functions_check.html).

## Ungrouped Functions

### Function `arglist()`

The main purpose of the package "kwb.utils" is to provide functions that allow
for writing cleaner and safer code. One main characteristic of clean code is
that it is free or at least almost free of duplication. This is a very hard goal 
to achieve! But we should try our best!

Have you ever seen something like the following?

```{r}
x <- seq(-pi, pi, pi/10)
par(mfrow = c(1, 2), mar = c(2, 2, 1, 1))
plot(x, sin(x), xlab = "x", ylab = "sin(x)", pch = 16, type = "b", cex = 0.5)
plot(x, cos(x), xlab = "x", ylab = "cos(x)", pch = 16, type = "b", cex = 0.5)
```

Do you spot the repetition? What if you want to change the plot character `pch`
or the `type` of the plot or the character extension factor `cex` in both
plots at the same time?

```{r}
# Define common arguments
common <- list(x = x, pch = 16, xlab = "x", type = "b", cex = 0.5)

# Call the plot function with the common arguments, extended by
# call-specific arguments
par(mfrow = c(1, 2), mar = c(2, 2, 1, 1))
do.call(plot, kwb.utils::arglist(common, y = sin(x), ylab = "sin(x)"))
do.call(plot, kwb.utils::arglist(common, y = cos(x), ylab = "cos(x)"))
```

The function `arglist()` merges the contents of one or more lists by allowing
to overwrite existing list elements. In the above example, the list of `common`
arguments is just extended. Elements in this list could also be overwritten,
such as the argument `cex` in the following:

```{r}
do.call(
  what = plot, 
  args = kwb.utils::arglist(common, y = sin(x), ylab = "sin(x)", cex = 1)
)
```

This pattern of calling a function with an extended argument list is so common 
that I wrote a function `callWith()` that allows to write the above shorter:

```{r}
kwb.utils::callWith(plot, common, y = sin(x), ylab = "sin(x)", cex = 1)
```

But this will be explained later...

### Function assignAll()
### Function assignArgumentDefaults()
### Function assignGlobally()
### Function assignPackageObjects()
### Function breakInSequence()
### Function callWith()
### Function callWithStringsAsFactors()
### Function catAndRun()
### Function catIf()
### Function catLines()
### Function catNewLineIf()
### Function checkForMissingColumns()
### Function clearConsole()
### Function cmdLinePath()
### Function colMaxima()
### Function colMinima()
### Function colNaNumbers()
### Function colStatisticOneFunction()
### Function colStatistics()
### Function columnDescriptor()
### Function columnToDate()
### Function columnwisePercentage()
### Function combineAlternatingly()
### Function compareDataFrames()
### Function compareSets()
### Function convertCsvFile()
### Function copyDirectoryStructure()
### Function copyListElements()
### Function countNaInColumn()
### Function countOrSum()
### Function createDirAndReturnPath()
### Function createDirectory()
### Function createMatrix()
### Function createPasswordFile()
### Function csvTextToDataFrame()
### Function defaultIfNA()
### Function defaultIfNULL()
### Function defaultIfZero()
### Function defaultWindowsProgramFolders()
### Function desktop()
### Function diffrows()
### Function DIN.A4()
### Function dropDim()
### Function dropUnusedFactorLevels()
### Function enlargeVector()
### Function excludeNULL()
### Function expandGrid()
### Function extendLimits()
### Function extractRowRanges()
### Function fileExtension()
### Function finishAndShowPdf()
### Function finishAndShowPdfIf()
### Function firstElement()
### Function firstPosixColumn()
### Function frequencyTable()
### Function fullySorted()
### Function generateKeyFile()
### Function get_homedir()
### Function getAttribute()
### Function getByPositiveOrNegativeIndex()
### Function getEvenNumbers()
### Function getFunctionName()
### Function getFunctionValueOrDefault()
### Function getGlobally()
### Function getKeywordPositions()
### Function getNamesOfObjectsInRDataFiles()
### Function getObjectFromRDataFile()
### Function getOddNumbers()
### Function getPassword()
### Function getPathsAndValuesFromRecursiveList()
### Function getTagNames()
### Function guessSeparator()
### Function headtail()
### Function hsAddMissingCols()
### Function hsChrToNum()
### Function hsDelEmptyCols()
### Function hsMatrixToListForm()
### Function hsMovingMean()
### Function hsOpenWindowsExplorer()
### Function hsPrepPdf()
### Function hsRenameColumns()
### Function hsResolve()
### Function hsRestoreAttributes()
### Function hsShell()
### Function hsShowPdf()
### Function hsStringToDate()
### Function hsStringToDouble()
### Function hsSystem()
### Function insertColumns()
### Function lastElement()
### Function limitToRange()
### Function linearCombination()
### Function listObjects()
### Function loadObject()
### Function makeUnique()
### Function mergeAll()
### Function mergeLists()
### Function mergeNamedArrays()
### Function moveColumnsToFront()
### Function movingSum()
### Function msgAvailableFields()
### Function mySystemTime()
### Function nameByElement()
### Function naToLastNonNa()
### Function noFactorDataFrame()
### Function objectSize()
### Function parallelNonNA()
### Function pasteColumns()
### Function pasteColumns0()
### Function percentage()
### Function percentageOfMaximum()
### Function percentageOfSum()
### Function posixColumnAtPosition()
### Function preparePdf()
### Function preparePdfIf()
### Function printIf()
### Function quotient()
### Function randomMatrix()
### Function randomValuesWithSum()
### Function rbindAll()
### Function readArglists()
### Function readCsvInputFile()
### Function readDictionaries()
### Function readDictionary()
### Function readPackageFile()
### Function recursiveNames()
### Function recycle()
### Function relativeCumulatedSum()
### Function removeAttributes()
### Function removeColumns()
### Function removeElements()
### Function removeEmptyColumns()
### Function renameAndSelect()
### Function renameColumns()
### Function resetRowNames()
### Function resolve()
### Function resolveAll()
### Function revertListAssignments()
### Function roundColumns()
### Function rowOrColumnwisePercentage()
### Function rowwisePercentage()
### Function rStylePath()
### Function runBatchfileInDirectory()
### Function runInDirectory()
### Function safeColumnBind()
### Function safeMerge()
### Function safePath()
### Function safeRowBind()
### Function safeRowBindAll()
### Function safeRowBindOfListElements()
### Function selectColumns()
### Function selectElements()
### Function setColumns()
### Function setLoaded()
### Function setMatrixColumns()
### Function sourceScripts()
### Function splitAlongDim()
### Function splitIntoFixSizedBlocks()
### Function startsToEnds()
### Function startsToRanges()
### Function stopIfNotMatrix()
### Function tableLookup()
### Function tempSubdirectory()
### Function toFactor()
### Function toFormula()
### Function toInches()
### Function toKeysAndValues()
### Function toLookupClass()
### Function toLookupList()
### Function toLookupTable()
### Function toNamedList()
### Function toPositiveIndices()
### Function underscoreToPercent()
### Function unmerge()
### Function user()
### Function warnIfEmpty()
### Function warningDeprecated()
### Function windowsPath()
### Function writeDictionary()
### Function writeText()
